---
title: "Week 1 Assignment"
author: "Samriti Malhotra"
date: "January 30, 2019"
output:
  pdf_document: default
  html_document: default
---
#Introduction
Week 1 assignment for DATA 607 is to subset the data provided by the UCI for Mushroom dataset located at below address:

https://archive.ics.uci.edu/ml/datasets/Mushroom

The actual dataset can be found here:

https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data

#About the Data

This data set includes descriptions of hypothetical samples corresponding to 23 species of gilled mushrooms in the Agaricus and Lepiota Family.Each species is identified as definitely edible, definitely poisonous, or of unknown edibility and not recommended. This latter class was combined with the poisonous one

#Data Dictionary

Attribute Information:

0. Type     : Edible e , Poisonous = p

1. cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s 

2. cap-surface: fibrous=f,grooves=g,scaly=y,smooth=s 

3. cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r, pink=p,purple=u,red=e,white=w,yellow=y 

4. bruises: bruises=t,no=f 

5. odor: almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s 

6. gill-attachment: attached=a,descending=d,free=f,notched=n 

7. gill-spacing: close=c,crowded=w,distant=d 

8. gill-size: broad=b,narrow=n 

9. gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e, white=w,yellow=y 

10. stalk-shape: enlarging=e,tapering=t 

11. stalk-root: bulbous=b,club=c,cup=u,equal=e, rhizomorphs=z,rooted=r,missing=? 

12. stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s 

13. stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s 

14. stalk-color-above-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y 

15. stalk-color-below-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y 

16. veil-type: partial=p,universal=u 

17. veil-color: brown=n,orange=o,white=w,yellow=y 

18. ring-number: none=n,one=o,two=t 

19. ring-type: cobwebby=c,evanescent=e,flaring=f,large=l, none=n,pendant=p,sheathing=s,zone=z 

20. spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r, orange=o,purple=u,white=w,yellow=y 

21. population: abundant=a,clustered=c,numerous=n, scattered=s,several=v,solitary=y 

22. habitat: grasses=g,leaves=l,meadows=m,paths=p, urban=u,waste=w,woods=d


#Problem Statement:-
Load data from given URL(https://archive.ics.uci.edu/ml/datasets/Mushroom) into R , subset and create a new Data frame selecting few columns from original including 1st column. Provide meangingfull name to columns Headers, and also update the values of each column based on the data dictionary values. 

##Loading Libraries
First step to load the necessary libraries required for this assignment.
```{r}
library(stringr)
library(XML)
library(maps)
library(httr)
library(dplyr)
```


## R-code

```{r}
mushroom_table <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data",header=FALSE , sep="," )

mushrooms <- as.data.frame(mushroom_table)


# subsetting the data to create a a new data frame with 5 columns 
mush_subset <- subset(mushrooms, select=c(1,2,3,4,6))

#providing columns meaningfull names
colnames(mush_subset) <- c("Type","Shape","Surface","Color","Odor")

#Changing the column values with meaningfull values based on conditional statements
mush_subset$Type <- ifelse(str_detect(mush_subset$Type, "e") == TRUE, "Edible", "Poisonous")
ELSE <- TRUE

##Pipe operator %>% allows you to pipe the output of one function to the input of another functione. It works from left to right
##mutate function will add new columns to dataframe
## with function hlps in constructing an environemnt from data, possible modifying(a copy of) the orignial data.e,g.  with(data,expression,....)
#bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s
mush_subset <- mush_subset %>% mutate(.,Shape1 = with(.,case_when(
  (mush_subset$Shape == "x") ~ "convex",
  (mush_subset$Shape == "b") ~ "bell",
  (mush_subset$Shape == "c") ~ "conical",
  (mush_subset$Shape == "k") ~ "knobbed",
  (mush_subset$Shape == "f") ~ "flat",
  ELSE ~ "sunken"
)))

#fibrous=f,grooves=g,scaly=y,smooth=s
mush_subset <- mush_subset %>% mutate(.,Surface = with(.,case_when(
  (mush_subset$Surface == "f") ~ "fibrous",
  (mush_subset$Surface == "g") ~ "grooves",
  (mush_subset$Surface == "y") ~ "scaly",
  ELSE ~ "smooth"
)))

#brown=n,buff=b,cinnamon=c,gray=g,green=r, pink=p,purple=u,red=e,white=w,yellow=y
mush_subset <- mush_subset %>% mutate(.,Color = with(.,case_when(
  (mush_subset$Color == "n") ~ "brown",
  (mush_subset$Color == "b") ~ "buff",
  (mush_subset$Color == "c") ~ "cinnamon",
  (mush_subset$Color == "g") ~ "gray",
  (mush_subset$Color == "r") ~ "green",
  (mush_subset$Color == "p") ~ "pink",
  (mush_subset$Color == "u") ~ "purple",
  (mush_subset$Color == "e") ~ "red",
  (mush_subset$Color == "w") ~ "white",
  ELSE ~ "yellow"
)))

#almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s
mush_subset <- mush_subset %>% mutate(.,Odor = with(.,case_when(
  (mush_subset$Odor == "a") ~ "almond",
  (mush_subset$Odor == "l") ~ "anise",
  (mush_subset$Odor == "c") ~ "creosote",
  (mush_subset$Odor == "y") ~ "fishy",
  (mush_subset$Odor == "f") ~ "foul",
  (mush_subset$Odor == "m") ~ "musty",
  (mush_subset$Odor == "p") ~ "pungent",
  (mush_subset$Odor == "n") ~ "none",
  ELSE ~ "spicy"
  
  
)))


```






##Summary
The data frame after subsetting, has been given meaningfull name and updating column values is as below.
```{r}
head(mush_subset,n=20)
```

