---
title: "Week 3 Assignment"
author: "Samriti Malhotra"
date: "February 17, 2019"
output:
  html_document: default
  pdf_document: default
partner: Vishal Arora
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction#
Using regular expression and string manipulation to extract relevant information from structured, un-structed text data using R. The below few examples from Week3 assignment, shows how we can use regular expression to extract strings, on which we can then apply R packages like <B><i>stringr</i></B> / <b><i>concatenation</b></i> / <b><i>BaseR</b></i> etc.. to manipulate the strings and construct something meangingfull.

```{r}
library(stringr)
library(concatenate)
```


## Question 3:- ##
Copy the introductory example. The vector name stores the extracted names.

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.
(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).
(c) Construct a logical vector indicating whether a character has a second name.

### Answer 3:-###
```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer55536420Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
phoneNum <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?(\\d{3})(-| )?(\\d{4})"))

# a) 
name1 <- str_replace(name ,pattern = "C. " , replacement = "")
name1 <- str_replace(name1 ,pattern = "," , replacement = "")
unlist(str_extract_all(name1, "[[:alpha:],]{2,}[:space:]+[:alpha:]{2,}"))

# b)
str_detect(name,"[[:alpha:]]{2,3}[\\.]")

# c) 
str_detect(name,"[:SPACE:][[:alpha:]]{1,}[\\.][:SPACE:]")
```

## Question 4:- ##
Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.<br />
(a) [0-9]+\\$ <br />
(b) \\b[a-z]{1,4}\\b <br />
(c) .*?\\.txt$ <br />
(d) \\d{2}/\\d{2}/\\d{4} <br />
(e) <(.+?)>.+?</\\1>

### Answer 4 :-###
```{r}
# (i)
str_4i <- c("ADGYRE67575$", "1245638ABCV$","123BAC45")
str_detect(str_4i,"[0-9]+\\$" )

# (ii)
str_4ii <- c( "test in the week","Cat", "1234")
str_detect(str_4ii,"\\b[a-z]{1,4}\\b")

# (iii)
strPattern <- ".*?\\.txt$"
str_4iii <- c("Assingment.txt","assignment")
str_detect(str_4iii,strPattern)

# (iv)
strPatt <- "\\d{2}/\\d{2}/\\d{4}"
str_4iv <-c("2/2/1977","02/02/1977","1977/03/29","03/29/86")
str_detect(str_4iv,strPatt)

# (v)
strPatt <- "<(.+?)>.+?</\\1>"
str_4v <- c("<h1> this is an paragraph header</h1>","<p>test</pt>","<test>b</test")
str_detect(str_4v,strPatt)
```


## Question 9:- ##
The following code hides a secret message. Crack it with R and regular expressions.
Hint: Some of the characters are more revealing than others! The code snippet is also
available in the materials at www.r-datacollection.com.
clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr


###Answer 9 :- ###
```{r}
secret_mesg <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")

secret_msg <- unlist(str_extract_all(secret_mesg,"[[:upper:].]{1,}"))
secret_msg <- cc(secret_msg)
secret_msg <- str_replace_all(secret_msg,",","")
secret_msg <- str_replace_all(secret_msg,"\\."," ")
secret_msg
```

#Summary#
During these examples we learnt various techniques of regular expression(back referencing, fixed, {}repeators, [:alpha/space:]) .
Also used various methods of stringr , concetanation packages to manipulate the strings.
