---
title: "Project2_C"
author:
- name: Samriti Malhotra
- name: Vishal Arora
date: "3/10/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction :-
As part of project 2, this is the third messy data which needs to cleaned and tranformed to be ready to do the analysis.


##Problem Statment :-
This data set contains wine ratings of different varieties which are from different countries,the data has lot of blank values for wine points/country which needs to be cleaned and the data is mismanaged which needs to first sorted and arranged in way, before any meaningfull analysis can be done.


##Solution :-
We are using below libraries in our quest to resolve the above problem:-
*knitr*    
*tidyr*     
*dplyr*      
*kableExtra*      
*ggplot2*       
```{r include=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

####Steps performed on cleaning, tranforming and doing Analysis on the data    
  *1) Reading the data from local directory using read.csv function.*      
  *2) Setting the blank data to NA and then we selected the relevant columns for our cleaning up and analysis using select function from tidy.*      
```{r}
raw_df <- read.csv('winemag-data-130k-v2.csv', stringsAsFactors=F)
raw_df[raw_df==""] <- NA

wine_data <- select(raw_df , 2,5,6,7,13,14)
kable(head(wine_data)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed","responsive"),full_width   = F,position = "left",font_size = 12) %>%
  row_spec(0, background ="gray")
```
  *3) Then we removed the records which had Country as NA, using drop_na from tidy library.*       
  *4) Next step we are calculating the average points per country wise , using group_by, summarize and applying mean on points for that country all these             functions are from dplyr library.*      
  *5) we used glinmpse function from tidyr to have a snapshot of our final data set.*

```{r}
wine_data <- wine_data %>%   drop_na(country)
wine_data_variety <- wine_data %>%
  group_by(country ) %>%
  summarise(avg_points = round(mean(points)))

kable(head(wine_data_variety)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed","responsive"),full_width   = F,position = "left",font_size = 12) %>%
  row_spec(0, background ="gray")

```

   *6) Using the ggplot2 library we plot the average points against Country names to do our analysis.*    

```{r inlcude=FALSE}
wine_data_variety %>% 
  ggplot(aes(x=country, y=avg_points)) + 
  geom_point() + geom_line() + 
  labs(title = "Average Points  per Country", x="Country", y="Points") +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

##Summary 
Looking at the plotted graph we can say that the wines from England has been adjudged the best with 92 average wine tasting points, closely followed by Austria, Germany and India with 90 points. whereas Egypt,Peru and Ukraine wines have the least wine tasting average points.




